---
title: "Project"
author: "Shawn Riemer"
date: "11/11/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(corrplot)
library(corrr)
```

## Read in Data

```{r}
country <- read_csv("data/Country.csv", col_types = cols())
ind <- read_csv("data/Indicators.csv", col_types = cols())
```

```{r}
# Each observation is uniquely identified by CountryName/CountryCode, IndicatorName/IndicatorCode, Year
# ind %>% str()
# ind$Year %>% unique()

names <- read_tsv("world-country-names.tsv")
data_names <- names$name
ind_names <- ind$CountryName %>% unique()
overlap_names <- intersect(ind_names, data_names)
overlap_names %>% length()

ind <- ind %>% 
  mutate(
    IndicatorName = str_replace_all(IndicatorName, " ", "_")
  ) 
ind <- ind %>% filter(
    CountryName != "Channel Islands" # remoce bc it has an entire missing year
  ) %>% filter(
    CountryName %in% overlap_names
  )
```

```{r}
# pivot wider to add missing values
ind2 <- ind %>%
  select(CountryName, IndicatorCode, Year, Value) %>%
  pivot_wider(names_from = IndicatorCode,
              values_from = Value,
              values_fill = NA)

# foo <- ind2 %>% group_by(CountryName) %>% 
#   summarise(
#     n = n()
#   )
# foo$n %>% unique()

# foo %>% filter(n != 56)
```


```{r}
p_missing_ <- function(d) {
  mvt <- table(is.na(d))
  mvt[2] / (mvt[1] + mvt[2])
}

p_missing_(ind2)

# ind$IndicatorCode %>% unique() %>% length()
# 1344
```

Number of missing values for each indicator
```{r}
max_percent_missing <- 0.2

# Filter to only have indicators that have less than max_percent_missing percent of missing values
sapply(ind2, function(x) sum(is.na (x))) %>% 
  as.data.frame() %>% 
  select(., everything()) %>% 
  mutate(
    p_missing = . / 13823
  ) %>% 
  arrange(-desc(.)) %>% 
  filter(p_missing < max_percent_missing)

(foo <- sapply(ind2, function(x) sum(is.na (x))) %>% 
  as.data.frame() %>% 
  select(., everything()) %>% 
  mutate(
    p_missing = . / 13823
  ) %>% 
  arrange(-desc(.)) %>% 
  filter(p_missing < 0.5))
         
filtered_indicators <- foo %>% row.names()
filtered_indicators <- filtered_indicators[filtered_indicators != "Year"]
filtered_indicators %>% length()
```

Number of missing values per country
```{r}
total_num_data_ideally <- (ind2 %>% colnames() %>% length() - 2) * 56 # CountryName and Year don't count
max_percent_missing <- 0.1

# IF JOO
# total_num_data_ideally <- total_num_data_ideally - (filtered_indicators %>% length()) * 56

(good_countries <- ind2 %>% 
  group_by(CountryName) %>% 
  select(-Year) %>% 
  # select(filtered_indicators) %>%
  summarise_all(funs(sum(is.na(.)))) %>% 
  mutate(
    p_missing = select(., -CountryName) %>% rowSums() / total_num_data_ideally
  ) %>% select(
    CountryName,
    p_missing
  ) %>% arrange(desc(p_missing)) %>% 
  filter(p_missing < 0.7))

ind3 <- ind2 %>% 
  filter(CountryName %in% good_countries$CountryName)

(chosen_indicators_ <- sapply(ind3, function(x) sum(is.na (x))) %>% 
  as.data.frame() %>% 
  select(., everything()) %>% 
  mutate(
    p_missing = . / 13823
  ) %>% 
  arrange(-desc(.)) %>% 
  filter(p_missing < max_percent_missing))
chosen_indicators <- chosen_indicators_ %>% row.names()

ind4 <- ind3 %>% 
  select(CountryName, Year, chosen_indicators)

p_missing_(ind4)

ind4$`GDP_at_market_prices_(current_US$)` %>% is.na() %>% sum() / ind4 %>% nrow() #
ind4$`GDP_at_market_prices_(constant_2005_US$)` %>% is.na() %>% sum() / ind4 %>% nrow()
ind4$`GDP_per_capita_(constant_2005_US$)` %>% is.na() %>% sum() / ind4 %>% nrow()
ind4$`GDP_per_capita_(current_US$)` %>% is.na() %>% sum() / ind4 %>% nrow() # 0.09963284 missing <- good to predict
ind4$`GDP_growth_(annual_%)` %>% is.na() %>% sum() / ind4 %>% nrow()
ind4$`GDP_per_capita_growth_(annual_%)` %>% is.na() %>% sum() / ind4 %>% nrow()
```

We want to have some fraction of countries we remove and some fraction of indicators we remove such that the total number of missing values is low.

```{r}
ind4$CountryName %>% unique() %>% length()
# prcomp(ind4)

# ind4 %>% pivot_wider()

ind5 <- ind4 %>% group_by(CountryName) %>% 
  select(-Year) %>%
  summarise_all(
    funs(mean(., na.rm = TRUE))
  )
ind5
```


```{r}
for (i in 2:ncol(ind5)) {
  ind5[is.na(ind5[,i]), i] <- lapply(ind5[,i], mean, na.rm=TRUE)
}
```

```{r}
ind5

code_to_name <- function(code) {
  uh <- ind %>% select(IndicatorName, IndicatorCode) %>% filter(IndicatorCode == code) %>% unique()
  uh$IndicatorName
}

name_to_code <- function(name) {
  uh <- ind %>% select(IndicatorName, IndicatorCode) %>% filter(IndicatorName == name) %>% unique()
  uh$IndicatorCode
}

gdp_per_capita_current_US <- ind5 %>%
  select(name_to_code("GDP_per_capita_(current_US$)"))
predictors <- ind5 %>%
  select(-name_to_code("GDP_per_capita_(current_US$)"),
         -"CountryName") 

pc3 <- prcomp(predictors,scale=TRUE, rank. = 2)
km3 <- kmeans(pc3$x,centers=5, nstart=10)

pcs <- pc3$x %>% data.frame()

km3$cluster %>% length()

foo <- cbind(ind5, pcs) %>% select(
  CountryName, PC1, PC2, everything()
)

foo$Cluster = km3$cluster

foo %>% 
  select(
    contains("GDP"),
    contains("GNI")
  )

foo <- foo %>% select(
  CountryName, Cluster, PC1, PC2, everything()
)

pairs(foo[,5:9], col=foo$Cluster)

foo[,5:142] %>% scale() %>% data.frame()

# foo[,5:142] %>% dist() %>% hclust() %>% plot(labels=FALSE)

foo

foo[,5:142] %>% scale() %>% data.frame() %>% dist() %>% hclust() %>% plot(labels=FALSE)
# foo[,5:142] %>% scale() %>% data.frame() %>% dist() %>% hclust() %>% plot()

my_colcode <- foo[,5:9] %>% colnames()
my_colnames <- foo[,5:9] %>% colnames() %>% code_to_name()
```

```{python}
codes = r.my_colcode
names = r.my_colnames

print("| Code | Name |")
print("|-|-|")
for c,n in zip(codes, names):
  print(f"| {c} | {n} | ")
```
| Code | Name |
|-|-|
| SP.ADO.TFRT | Adolescent_fertility_rate_(births_per_1,000_women_ages_15-19)| 
| SP.POP.DPND | Age_dependency_ratio_(%_of_working-age_population)| 
| SP.POP.DPND.OL | Age_dependency_ratio,_old_(%_of_working-age_population)| 
| SP.POP.DPND.YG | Age_dependency_ratio,_young_(%_of_working-age_population)| 
| SP.POP.65UP.TO.ZS | Population_ages_65_and_above_(%_of_total)| 

```{r}
# foo
# predictors
y <- gdp_per_capita_current_US
x <- prcomp(predictors,scale=TRUE, rank. = 5)$x %>% data.frame()

goo <- cbind(y, x) %>% rename(
  gdp_usd = `GDP_per_capita_(current_US$)`
)

model <- lm(gdp_usd ~ PC1 + PC2 + PC3 + PC4 + PC5, goo)
model %>% summary()
```





















